### 2023-04-16 Presentation.Rmd

---
title: "Presentation"
author: "Virginia Baskin"
date: "4/13/2023"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("shiny_plot_setup.R", echo=FALSE)
#show_col_types = FALSE
```

## Shiny Presentation

This R Markdown presentation is made interactive using Shiny. The viewers of the presentation can change the assumptions underlying what's presented and see the results immediately.

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Interactive Plot

```{r eruptions}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),

  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")

  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## R Output

```{r cars}
summary(cars)
```

## Data 
- Problem and Importance

## Data Introduction 
- bullet points 
- bullet points 
- bullet points 
- dude our data was all categorical so when i tell u it was difficult to plot

## Supplementary Datasets

## Exploration

- plot of citations by race over time  
- plot of frequency of citation versus hour in day  
- mph over speed limit by year

## Pyramid 
```{r, message = FALSE}
source("subscripts/citation_pyramid.R", local = TRUE, print.eval = FALSE, echo= FALSE, verbose= FALSE)
pyramid_cites()
```

## Violin
```{r, message = FALSE}
source("subscripts/violin_speed.R")
violin_speed()
```

--- 
## Killer Plot

## Shiny

---

#Race Line Graph by Year: line_ry()
#waffle plot: waffle()