### 2023-04-18 Final Presentation.Rmd
update

---
title: "Final Presentation"
author: "Siddhi Narayan, Grace Wang, Virginia Baskin"
date: "2023-04-17"
output: 
  ioslides_presentation:
    smaller: true
    widescreen: true
    css: style.css
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 9, fig.height = 5, out.width = "100%")
```

# Introduction

## Primary Dataset: Citations in HTX
Insert description

## Secondary Datasets
* Census Data
* GIS/Beats Data

## Exploration
- bullet 
- point

# Distribution Over Time

## Citation Count by Race over the Years {.smaller}
```{r Race/Year Line, message=FALSE}
source("subscripts/line_race_year.R", local=TRUE, print.eval=TRUE)
line_ry()
```

## Citation Frequency for Time of Day {.smaller}
```{r Time of Day Freq, message=FALSE}
source("subscripts/line_timeofcites.R", local=TRUE, print.eval=TRUE)
time_line()
```

# Distribution by Demographic

## Citation by Race {.smaller}
```{r Waffle, message=FALSE}
source("subscripts/waffle.R", local=TRUE, print.eval=TRUE)
waffle()
```

## MPH over Speed Limit for Race & Gender {.smaller}
```{r Violin, message=FALSE}
source("subscripts/violin_speed.R", local=TRUE, print.eval=TRUE)
violin_speed()
```

# Distribution by Location
* Plot - Citation by Sex by District 
* Table - Citations Per Square Mile

## Citation by Sex by District {.smaller}
```{r Pyramid, message = FALSE}
source("subscripts/citation_pyramid.R", local = TRUE, print.eval = TRUE)
pyramid_cites()
```

## Citations Per Square Mile {.smaller}
* table
* table 
* table

## Killer Plot
<iframe src="http://127.0.0.1:5290" width="100%" height="500px"></iframe>

# Analysis

## Houston Racial Breakdown vs Citations Racial Breakdown {.smaller}
table

## Citation Type Racial Breakdown {.smaller}
table

## Failure to Establish Financial Responsibility {.smaller}
table

## Average Speeding Amounts {.smaller}
table

# Conclusion

## Conclusions
conclusions