### 2023-04-18 Final Presentation.Rmd
slay

---
title: "Final Presentation"
author: "Siddhi Narayan, Grace Wang, Virginia Baskin"
date: "2023-04-17"
output: 
  ioslides_presentation:
    smaller: true
    widescreen: true
    css: style.css
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 9, fig.height = 5, out.width = "100%")
source("shiny_plot_setup.R", echo=FALSE)
source("Tables/Racial_Citation.R", echo = FALSE)
source("Tables/Racial_Breakdown.R", echo = FALSE)
source("Tables/FinResp.R", echo = FALSE)
source("Tables/average_speeding_amt.R", echo = FALSE)
source("Tables/csqm.R", echo = FALSE)
```

# Introduction

## Primary Dataset: Citations in HTX
Insert description

## Secondary Datasets
* Census Data
* GIS/Beats Data

## Exploration
- bullet 
- point

# Distribution Over Time

## Citation Count by Race over the Years {.smaller}
```{r Race/Year Line, message=FALSE}
source("subscripts/line_race_year.R", local=TRUE, print.eval=TRUE)
line_ry()
```

## Citation Frequency for Time of Day {.smaller}
```{r Time of Day Freq, message=FALSE}
source("subscripts/line_timeofcites.R", local=TRUE, print.eval=TRUE)
time_line()
```

# Distribution by Demographic

## Citation by Race {.smaller}
```{r Waffle, message=FALSE}
source("subscripts/waffle.R", local=TRUE, print.eval=TRUE)
waffle()
```

## MPH over Speed Limit for Race & Gender {.smaller}
```{r Violin, message=FALSE}
source("subscripts/violin_speed.R", local=TRUE, print.eval=TRUE)
violin_speed()
```

# Distribution by Location
* Plot - Citation by Sex by District 
* Table - Citations Per Square Mile

## Citation by Sex by District {.smaller}
```{r Pyramid, message = FALSE}
source("subscripts/citation_pyramid.R", local = TRUE, print.eval = TRUE)
pyramid_cites()
```

## Citations Per Square Mile {.smaller}
Top 5 Beats ranked by citations/square mile:
```{r csqm_top, echo = FALSE, message = FALSE}
knitr::kable(csqm_top)
```
##
Bottom 5 Beats ranked by citations/square mile
```{r csqm_bottom, echo = FALSE, message = FALSE}
knitr::kable(csqm_bottom)
```
##
Statistics on number of citations/square mile:

```{r csqm_summry, echo = FALSE, message = FALSE}
knitr::kable(cqm_summary)
```

## Killer Plot
<iframe src="http://127.0.0.1:5290" width="100%" height="500px"></iframe>

# Analysis

## Houston Racial Breakdown vs Citations Racial Breakdown {.smaller}
```{r racial_citation_breakdown, echo = FALSE, message = FALSE}
knitr::kable(htx)
```

## Citation Type Racial Breakdown {.smaller}
Speeding Citations by Race:
```{r speeding_race, echo = FALSE, message = FALSE}
knitr::kable(speeding_race)
```
##
Invalid License Citations by Race:
```{r license_race, echo = FALSE, message = FALSE}
knitr::kable(license_race)
```

##
Failure to Establish Financial Responsibility Citations by Race:
```{r financial_race, echo = FALSE, message = FALSE}
knitr::kable(financial_race)
```

##
Seat Belt Citations by Race:
```{r seat_belt_race, echo = FALSE, message = FALSE}

knitr::kable(seat_belt_race)
```

##
Running a Red Light/Stop Sign Citations by Race:
```{r stop_light_race, message = FALSE}
knitr::kable(stop_light_race)
```

## Failure to Establish Financial Responsibility {.smaller}
```{r fin_resp, echo = FALSE, message = FALSE}
knitr::kable(all_perc)
```

## Average Speeding Amounts {.smaller}
All People (baseline):
```{r all_speed, echo = FALSE, message = FALSE}
knitr::kable(all_speed)
```
##
By Race:
```{r race_speed, echo = FALSE, message = FALSE}
knitr::kable(race_speed)
```
##
By Gender:
```{r gender_speed, echo = FALSE, message = FALSE}
knitr::kable(gender_speed)
```
##
```{r race_gender_speed, echo = FALSE, message = FALSE}
library(kableExtra)
kable(race_gender_speed) %>% kable

# Conclusion

## Conclusions
conclusions